Cochain level May--Steenrod operations

### 2021-08-30 sec/effective.tex
Update references and split-off acks
Use terminology cup-(r,i) products
Improvements after Greg's proofreading

In this section we construct explicit May--Steenrod structures on three well know combinatorial $E_\infty$-operads: the Barratt--Eccles operad $\mathcal E$ \cite{berger2004combinatorial}, the surjection operad $\mathcal X$ \cite{mcclure2003multivariable}, and the operad $U(\mathcal M)$ associated to the finitely presented $E_\infty$-prop $\mathcal M$ introduced in \cite{medina2020prop1}.
We also define natural and effective May--Steenrod structure on the normalized cochains of any simplicial or cubical set using that these are algebras over the operad $U(\mathcal M)$.

@@

\begin{figure}
	\begin{tikzcd}
	& & & U(\mathcal M) \arrow[dddr, bend left=10, "\phi^\triangle"'] \arrow[dddrr, bend left=20, "\phi^\square"'] & & \\
	& & \mathcal X \arrow[ru, "SL"] & & & \\
	& \mathcal E \arrow[ur, "TR"] & & & & \\
	\mathcal W \arrow[rrrr, "\psi^\triangle", bend right=10] \arrow[rrrrr, "\psi^\square", bend right=20] \arrow[ur, "\psi_{\mathcal E}"] \arrow[uurr, "\psi_{\mathcal X}"', bend right=35, pos=.6] \arrow[uuurrr, "\psi_{U(\mathcal M)}"', bend right=45, pos=0.6] & & & & \End_{N^\bullet(X)} & \End_{N^\bullet(X)}
	\end{tikzcd}
	\caption{Summary of effective constructions: May--Steenrod structures on the Barratt--Eccles $\mathcal E$, surjection $\mathcal X$, and $U(\mathcal M)$ operads, and natural May--Steenrod structures on the normalized chains of simplicial and cubical sets. We remark that the maps $TR$ and $SL$ required different sign conventions.}
	\label{fig: bigsummary}
\end{figure}

\subsection{Barratt--Eccles operad} In this subsection we effectively describe a May--Steenrod structure on the Barratt--Eccles operad.

We begin by reviewing the $\mathrm{S}$-module structure underlying the Barratt--Eccles operad and, since will not use it in this work, refer to \cite{berger2004combinatorial} for a description of if composition structure.
For a non-negative integer $r$ define the simplicial set $E(\mathrm S_r)$ by
\begin{equation} \label{eq: milnor model of symmetric}
\begin{split}
E(\mathrm S_r)_n &= \{ (\sigma_0, \dots, \sigma_n)\ |\ \sigma_i \in \mathrm{S}_r\}, \\
d_i(\sigma_0, \dots, \sigma_n) &= (\sigma_0, \dots, \widehat{\sigma}_i, \dots, \sigma_n), \\
s_i(\sigma_0, \dots, \sigma_n) &= (\sigma_0, \dots, \sigma_i, \sigma_i, \dots, \sigma_n) \\
\end{split}
\end{equation}
with a left $\mathrm S_r$-action given by
\begin{equation*}
\sigma (\sigma_0, \dots, \sigma_n) = (\sigma \sigma_0, \dots, \sigma \sigma_n).
\end{equation*}
The chain complex resulting from applying the functor of normalized integral chains to it is denoted $\mathcal E(r)$, which corresponds to the arity $r$ part of the Barratt--Eccles operad.

@@

\begin{proof}
	Since $\mathcal X$ is an $E_\infty$-operad, we simply need to prove that the $\Z[\mathrm{C}_r]$-linear map
	\begin{equation*}
	\psi_{\mathcal E}(r) \colon \mathcal W(r) \to \mathcal E(r)
	\end{equation*}
	is a chain map for every $r \geq 0$.
	We simplify notation and write $\psi$ instead of $\psi_{\mathcal E}(r)$.
	To show $\psi$ is a chain map we proceed by induction.
	Notice that
	\begin{equation*}
	\psi(\partial e_0) = 0 = \partial \psi(e_0)
	\end{equation*}
	and assume $\psi(\partial e_{k-1}) = \partial \psi(e_{k-1})$.
	If $k = 2n$ we have
	\begin{align*}
	\partial \psi(e_{2n}) & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^{r_1}, \rho^{r_1+1}, \dots, \rho^{r_n}, \rho^{r_n+1} \big) \\ & =
	\partial \sum_{r_2, \dots, r_n} \sum_{r_1 = 0}^{p-1}
	\big(\rho^0, \rho^{r_1} \, (\rho^0, \rho^{1}, \dots, \rho^{r_n-r_1}, \rho^{r_n - r_1 +1}) \big) \\ & =
	\partial \sum_{r_2, \dots, r_n}
	\big(\rho^0, N\, (\rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n + 1}) \big) \\ & =
	\sum_{r_2, \dots, r_n}
	N\, \big( \rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n + 1} \big) \\ & -
	\sum_{r_2, \dots, r_n}
	\big(\rho^0, \partial \, N \, (\rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n+1}) \big) \\ & =
	N \psi(e_{2n-1}) - (\rho^0, \partial N \psi (e_{2n-1})) \\ & =
	\psi(N e_{2n-1}) - (\rho^0, \psi (\partial N e_{2n-1})) \\ & =
	\psi(\partial e_{2n}) - (\rho^0, \psi (\partial^2 e_{2n})) \\ & =
	\psi(\partial e_{2n}).
	\end{align*}
	If $k = 2n+1$ we have
	\begin{align*}
	\partial \psi(e_{2n+1}) & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^1, \rho^{r_1}, \rho^{r_1+1}, \dots, \rho^{r_n}, \rho^{r_n+1} \big) \\ & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^{1} \, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, T\, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	\sum_{r_1, \dots, r_n}
	T \, \big( \rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n} \big) \\ & -
	\sum_{r_1, \dots, r_n}
	\big(\rho^0, \partial \, T \, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	T \psi(e_{2n}) - (\rho^0, \partial T \psi (e_{2n})) \\ & =
	\psi(T e_{2n}) - (\rho^0, \psi (\partial T e_{2n})) \\ & =
	\psi(\partial e_{2n+1}) - (\rho^0, \psi (\partial^2 e_{2n+1})) \\ & =
	\psi(\partial e_{2n+1})
	\end{align*}
	where for the third equality we used that for any $r_1, \dots, r_n$
	\begin{equation*}
	(\rho^0, \rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) = 0.
	\end{equation*}
	This map is a quasi-isomorphism since both complexes have the homology of a point and $\psi(e_0)$ represents a generator of the homology.
\end{proof}

@@

\begin{remark}
	The natural construction \eqref{eq: milnor model of symmetric} is defined for any group, in particular, for finite cyclic groups, and the inclusion $\iota \colon \mathrm C_r \to \mathrm S_r$ induces both; a simplicial inclusion $E(\iota) \colon E(\mathrm C_r) \to E(\mathrm S_r)$ and one of $\mathrm C$-modules
	\begin{equation*}
	N_\bullet E(\iota) \colon N_\bullet E(\mathrm C) \to N_\bullet E(\mathrm S) = \mathcal E.
	\end{equation*}
	We remark that the image of our map $\psi_{\mathcal E}$ lies in the subcomplex $N_\bullet E(\mathrm C)$, so the map $\psi_{\mathcal E}$ factors as
	\begin{equation*}
	\psi_{\mathcal E} \colon
	\begin{tikzcd}
	\mathcal W \arrow[r] &[-3pt] N_\bullet E(\mathrm C) \arrow[r, "N_\bullet E(\iota)\,"] &[5pt] \mathcal E.
	\end{tikzcd}
	\end{equation*}
\end{remark}

@@

In this subsection we introduce a May--Steenrod structures on the surjection operad.
There are two widely used sign conventions for this operad respectively due to McClure--Smith \cite{mcclure2003multivariable} and Berger--Fresse \cite{berger2004combinatorial}.
Using the Berger--Fresse convention, we can define a May--Steenrod structure on the surjection operad by composing the map $\psi_{\mathcal E}$ with the \textit{table reduction} quasi-isomorphism $\mathcal E \to \mathcal X$ introduced in Section 1.3 of \cite{berger2004combinatorial}.
We define an May--Steenrod structure on the surjection operad in a convention independent way which recovers the table reduction May--Steenrod structure in the Berger--Fresse case.

Let us start by recalling the definition of the surjection operad.
For a non-negative integer $r$ let $\mathcal X(r)_n$ be the free abelian group generated by all functions from $\{1, \dots, n+r\}$ to $\{1, \dots, r\}$ modulo the subgroup generated by degenerate functions, i.e., those which are either non-surjective or have a pair of equal consecutive values.
We only describe the McClure--Smith convention since we refer to it in subsequent sections.
The boundary map and symmetric action in this case is defined using the Koszul convention regarding a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\} $ as the top dimensional generator in the chain complex
\begin{equation} \label{eq: surjection boundary}
\bigotimes_{i = 1}^r N_\bullet \big( \triangle^{s^{-1}(i)}; \Z \big).
\end{equation}
Explicitly, if we think of $s$ as a sequence of integers $\big( s(1), \dots, s(n+r) \big)$ the boundary of $s$ is the sum of sequences obtained by orderly removing one at the time the occurrences of $1$ with alternating signs, then those of $2$ with starting sign equal to that of the last removal of $1$, and so on.
Since we do not use the composition structure we refer to \cite{mcclure2003multivariable} for it.

Let us recall the chain contraction of $\mathcal X(r)$ onto $\mathcal X(r-1)$ used by McClure--Smith to prove $\mathcal X$ is an $E_\infty$-operad and whose introduction is credited to Benson \cite{benson1998rep}.
Let the maps $i,\ p$ and $s$ be defined on basis elements, represented by sequences, as follows: $i \colon \mathcal X(r-1) \to \mathcal X(r)$ places a $1$ at the beginning of the sequence and increases each of the original entries by $1$, $p \colon \mathcal X(r) \to \mathcal X(r-1)$ takes the sequence to $0$ unless it contains a single occurrence of $1$, in which case $p$ removes the $1$ and decreases each of the remaining entries by $1$, and finally, $s \colon \mathcal X(r) \to \mathcal X(r)$ places a 1 at the beginning of the sequence; if the sequence already begins
with a 1, then the new sequence is degenerate so $s$ takes it to $0$.
These maps satisfy
\begin{equation*}
p i = \mathrm{id}
\qquad \text{ and } \qquad
\mathrm{id} - i p = \partial s + s \partial.
\end{equation*}
The compositions $i^{r-1}$ and $s^{r-1}$ define a contraction of $\mathcal X(r)$ onto $\mathcal X(1) \cong \Z$ with homotopy
\begin{equation*}
h = s + i\, s\, p + \cdots + i^{r-1} s\, p^{r-1},
\end{equation*}
i.e., they satisfy
\begin{equation*}
p^{r-1} i^{r-1} = \mathrm{id}
\qquad \text{ and } \qquad
\partial h + h \partial = \mathrm{id} - i^{r-1}\,p^{r-1}.
\end{equation*}

@@

\begin{proof}
	Since $\mathcal X$ is an $E_\infty$-operad, we simply need to prove that the $\Z[\mathrm{C}_r]$-linear map
	\begin{equation*}
	\psi_{\mathcal X}(r) \colon \mathcal W(r) \to \mathcal X(r)
	\end{equation*}
	introduced in Definition \ref{def: Steenrod-Adem on surjection} is a chain map for every $r \geq 0$.
	We simplify notation and write $\psi$ instead of $\psi_{\mathcal X}(r)$.
	To show $\psi$ is a chain map we proceed by induction.
	Notice that
	\begin{equation*}
	\psi(\partial e_0) = 0 = \partial \psi(e_0)
	\end{equation*}
	and assume $\psi(\partial e_{n-1}) = \partial \psi(e_{n-1})$.
	For $n = 2m+1$ we have
	\begin{align*}
	\partial \psi(e_{2m+1})
	& =
	\partial\, h\, T\, \psi(e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) - i^{r-1} p^{r-1}\, \psi(e_{2m}) -
	h\, \partial\, T\, \psi(e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) -
	h\, T\, \psi(\partial\, e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) -
	h\, \psi(T\,N\, e_{2m-1}) \big) \\
	& =
	T \psi(e_{2m}).
	\end{align*}
	For $n = 2m$ the proof is analogous.
	The chain map $\psi$ is a quasi-isomorphism since both complexes have the homology of a point and $\psi(e_0) = (1, \dots, r)$ is a generator of the homology.
\end{proof}

@@

We will define a May--Steenrod structure $\psi_{U(\mathcal M)} \colon \mathcal W \to U(\mathcal M)$ by composing the May--Steenrod structure $\psi_{\mathcal X} \colon \mathcal W \to \mathcal X$ with a morphism $SL \colon \mathcal X \to U(\mathcal M)$ of $\mathrm{C}$-modules ($\mathrm{S}$-modules in fact) that we now define.
Given a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\}$ let $SL(s)$ be the element represented by the immersed connected $(1,r)$-graph

>>>>

In this section we construct explicit May--Steenrod structures on three well know combinatorial $E_\infty$-operads: the Barratt--Eccles operad $\mathcal E$ \cite{berger2004combinatorial}, the surjection operad $\mathcal X$ \cite{mcclure2003multivariable}, and the operad $U(\mathcal M)$ associated to the finitely presented $E_\infty$-prop $\mathcal M$ introduced in \cite{medina2020prop1}.
We also define natural and effective May--Steenrod structures on the normalized cochains of any simplicial or cubical set using that these are algebras over the operad $U(\mathcal M)$.

@@

\begin{figure}[ht]
	\begin{tikzcd}
	& & & U(\mathcal M) \arrow[dddr, bend left=10, "\phi^\triangle_X"'] \arrow[dddrr, bend left=20, "\phi^\square_Y"'] & & \\
	& & \mathcal X \arrow[ru, "SL"] & & & \\
	& \mathcal E \arrow[ur, "TR"] & & & & \\
	\mathcal W \arrow[rrrr, "\psi^\triangle_X", bend right=10] \arrow[rrrrr, "\psi^\square_Y", bend right=20] \arrow[ur, "\psi_{\mathcal E}"] \arrow[uurr, "\psi_{\mathcal X}"', bend right=35, pos=.6] \arrow[uuurrr, "\psi_{U(\mathcal M)}"', bend right=45, pos=0.6] & & & & \End_{N^\bullet_\triangle(X)} & \End_{N^\bullet_\square(Y)}
	\end{tikzcd}
	\caption{Summary of effective constructions: May--Steenrod structures on the Barratt--Eccles $\mathcal E$, surjection $\mathcal X$, and $U(\mathcal M)$ operads, and natural May--Steenrod structures on the normalized chains of simplicial and cubical sets. We remark that the maps $TR$ and $SL$ required different sign conventions.}
	\label{fig: bigsummary}
\end{figure}

\subsection{Barratt--Eccles operad} In this subsection we effectively describe a May--Steenrod structure on the Barratt--Eccles operad via explicit formulae.

We begin by reviewing the $\mathrm{S}$-module structure underlying the Barratt--Eccles operad and, since will not use it in this work, refer to \cite{berger2004combinatorial} for a description of if composition structure.
For a non-negative integer $r$ define the simplicial set $E(\mathrm S_r)$ by
\begin{equation} \label{eq: milnor model of symmetric}
\begin{split}
E(\mathrm S_r)_n &= \{ (\sigma_0, \dots, \sigma_n)\ |\ \sigma_i \in \mathrm{S}_r\}, \\
d_i(\sigma_0, \dots, \sigma_n) &= (\sigma_0, \dots, \widehat{\sigma}_i, \dots, \sigma_n), \\
s_i(\sigma_0, \dots, \sigma_n) &= (\sigma_0, \dots, \sigma_i, \sigma_i, \dots, \sigma_n) \\
\end{split}
\end{equation}
with a left $\mathrm S_r$-action given by
\begin{equation*}
\sigma (\sigma_0, \dots, \sigma_n) = (\sigma \sigma_0, \dots, \sigma \sigma_n).
\end{equation*}
The chain complex resulting from applying the functor of normalized integral chains to it is denoted $\mathcal E(r)$, which is the arity $r$ chain complex of the Barratt--Eccles operad.

@@

\begin{proof}
	Since $\mathcal E$ is an $E_\infty$-operad, we simply need to prove that the $\Z[\mathrm{C}_r]$-linear map
	\begin{equation*}
	\psi_{\mathcal E}(r) \colon \mathcal W(r) \to \mathcal E(r)
	\end{equation*}
	is a quasi-isomorphism for every $r \geq 0$.
	We simplify notation and write $\psi$ instead of $\psi_{\mathcal E}(r)$.
	To show $\psi$ is a chain map we proceed by induction.
	Notice that
	\begin{equation*}
	\psi(\partial e_0) = 0 = \partial \psi(e_0)
	\end{equation*}
	and assume $\psi(\partial e_{k-1}) = \partial \psi(e_{k-1})$.
	If $k = 2n$ we have
	\begin{align*}
	\partial \psi(e_{2n}) & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^{r_1}, \rho^{r_1+1}, \dots, \rho^{r_n}, \rho^{r_n+1} \big) \\ & =
	\partial \sum_{r_2, \dots, r_n} \sum_{r_1 = 0}^{p-1}
	\big(\rho^0, \rho^{r_1} \, (\rho^0, \rho^{1}, \dots, \rho^{r_n-r_1}, \rho^{r_n - r_1 +1}) \big) \\ & =
	\partial \sum_{r_2, \dots, r_n}
	\big(\rho^0, N\, (\rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n + 1}) \big) \\ & =
	\sum_{r_2, \dots, r_n}
	N\, \big( \rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n + 1} \big) \\ & -
	\sum_{r_2, \dots, r_n}
	\big(\rho^0, \partial \, N \, (\rho^{0}, \rho^{1}, \dots, \rho^{r_n}, \rho^{r_n+1}) \big) \\ & =
	N \psi(e_{2n-1}) - (\rho^0, \partial N \psi (e_{2n-1})) \\ & =
	\psi(N e_{2n-1}) - (\rho^0, \psi (\partial N e_{2n-1})) \\ & =
	\psi(\partial e_{2n}) - (\rho^0, \psi (\partial^2 e_{2n})) \\ & =
	\psi(\partial e_{2n}).
	\end{align*}
	If $k = 2n+1$ we have
	\begin{align*}
	\partial \psi(e_{2n+1}) & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^1, \rho^{r_1}, \rho^{r_1+1}, \dots, \rho^{r_n}, \rho^{r_n+1} \big) \\ & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, \rho^{1} \, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	\partial \sum_{r_1, \dots, r_n}
	\big(\rho^0, T\, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	\sum_{r_1, \dots, r_n}
	T \, \big( \rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n} \big) \\ & -
	\sum_{r_1, \dots, r_n}
	\big(\rho^0, \partial \, T \, (\rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) \big) \\ & =
	T \psi(e_{2n}) - (\rho^0, \partial T \psi (e_{2n})) \\ & =
	\psi(T e_{2n}) - (\rho^0, \psi (\partial T e_{2n})) \\ & =
	\psi(\partial e_{2n+1}) - (\rho^0, \psi (\partial^2 e_{2n+1})) \\ & =
	\psi(\partial e_{2n+1})
	\end{align*}
	where for the third equality we used that for any $r_1, \dots, r_n$
	\begin{equation*}
	(\rho^0, \rho^0, \rho^{r_1-1}, \rho^{r_1}, \dots, \rho^{r_n - 1}, \rho^{r_n}) = 0.
	\end{equation*}
	This map is a quasi-isomorphism since both complexes have the homology of a point and $\psi(e_0)$ represents a generator of the homology.
\end{proof}

@@

\begin{remark}
	The natural construction \eqref{eq: milnor model of symmetric} is defined for any group, in particular, for finite cyclic groups, and the inclusion $\iota \colon \mathrm C_r \to \mathrm S_r$ induces both a simplicial inclusion $E(\iota) \colon E(\mathrm C_r) \to E(\mathrm S_r)$ and one of $\mathrm C$-modules
	\begin{equation*}
	N_\bullet E(\iota) \colon N_\bullet E(\mathrm C) \to N_\bullet E(\mathrm S) = \mathcal E.
	\end{equation*}
	We remark that the image of our map $\psi_{\mathcal E}$ lies in the subcomplex $N_\bullet E(\mathrm C)$, so the map $\psi_{\mathcal E}$ factors as
	\begin{equation*}
	\psi_{\mathcal E} \colon
	\begin{tikzcd}
	\mathcal W \arrow[r] &[-3pt] N_\bullet E(\mathrm C) \arrow[r, "N_\bullet E(\iota)\,"] &[5pt] \mathcal E.
	\end{tikzcd}
	\end{equation*}
\end{remark}

@@

In this subsection we introduce a May--Steenrod structures on the surjection operad.
There are two widely used sign conventions for this operad respectively due to McClure--Smith \cite{mcclure2003multivariable} and Berger--Fresse \cite{berger2004combinatorial}.
Using the Berger--Fresse convention, we can define a May--Steenrod structure on the surjection operad by composing the map $\psi_{\mathcal E}$ with the \textit{table reduction} quasi-isomorphism $\mathcal E \to \mathcal X$ introduced in Section 1.3 of \cite{berger2004combinatorial}.
We define a May--Steenrod structure on the surjection operad in a convention independent way which recovers the table reduction May--Steenrod structure in the Berger--Fresse case.

Let us start by recalling the definition of the surjection operad.
For a non-negative integer $r$ let $\mathcal X(r)_n$ be the free abelian group generated by all functions from $\{1, \dots, n+r\}$ to $\{1, \dots, r\}$ modulo the subgroup generated by degenerate functions, i.e., those which are either non-surjective or have a pair of equal consecutive values.
We only describe the McClure--Smith convention since we refer to it in subsequent sections.
The boundary map and symmetric action in this case is defined using the Koszul convention regarding a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\} $ as the top dimensional generator in the chain complex
\begin{equation} \label{eq: surjection boundary}
\bigotimes_{i = 1}^r N_\bullet \big( \triangle^{s^{-1}(i)}; \Z \big).
\end{equation}
Explicitly, if we think of $s$ as a sequence of integers $\big( s(1), \dots, s(n+r) \big)$ the boundary of $s$ is the sum of sequences obtained by orderly removing one at time the occurrences of $1$ with alternating signs, then those of $2$ with starting sign equal to that of the last removal of $1$, and so on.
Since we do not use the composition structure we refer to \cite{mcclure2003multivariable} for it.

Let us recall the chain contraction of $\mathcal X(r)$ onto $\mathcal X(r-1)$ used by McClure--Smith to prove $\mathcal X$ is an $E_\infty$-operad and whose introduction is credited to Benson \cite{benson1998rep}.
Let the maps $i$, $p$, and $s$ be defined on basis elements, represented by sequences, as follows: $i \colon \mathcal X(r-1) \to \mathcal X(r)$ places a $1$ at the beginning of the sequence and increases each of the original entries by $1$, $p \colon \mathcal X(r) \to \mathcal X(r-1)$ takes the sequence to $0$ unless it contains a single occurrence of $1$, in which case $p$ removes the $1$ and decreases each of the remaining entries by $1$, and finally, $s \colon \mathcal X(r) \to \mathcal X(r)$ places a 1 at the beginning of the sequence; if the sequence already begins
with a 1, then the new sequence is degenerate so $s$ takes it to $0$.
These maps satisfy
\begin{equation*}
p i = \mathrm{id}
\qquad \text{ and } \qquad
\mathrm{id} - i p = \partial s + s \partial.
\end{equation*}
The compositions $i^{r-1}$ and $s^{r-1}$ define a contraction of $\mathcal X(r)$ onto $\mathcal X(1) \cong \Z$ with homotopy
\begin{equation*}
h = s + i\, s\, p + \cdots + i^{r-1} s\, p^{r-1},
\end{equation*}
i.e., they satisfy
\begin{equation*}
p^{r-1} i^{r-1} = \mathrm{id}
\qquad \text{ and } \qquad
\partial h + h \partial = \mathrm{id} - i^{r-1}\,p^{r-1}.
\end{equation*}

@@

\begin{proof}
	Since $\mathcal X$ is an $E_\infty$-operad, we simply need to prove that the $\Z[\mathrm{C}_r]$-linear map
	\begin{equation*}
	\psi_{\mathcal X}(r) \colon \mathcal W(r) \to \mathcal X(r)
	\end{equation*}
	introduced in Definition \ref{def: Steenrod-Adem on surjection} is a quasi-isomorphism for every $r \geq 0$.
	We simplify notation and write $\psi$ instead of $\psi_{\mathcal X}(r)$.
	To show $\psi$ is a chain map we proceed by induction.
	Notice that
	\begin{equation*}
	\psi(\partial e_0) = 0 = \partial \psi(e_0)
	\end{equation*}
	and assume $\psi(\partial e_{n-1}) = \partial \psi(e_{n-1})$.
	For $n = 2m+1$ we have
	\begin{align*}
	\partial \psi(e_{2m+1})
	& =
	\partial\, h\, T\, \psi(e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) - i^{r-1} p^{r-1}\, \psi(e_{2m}) -
	h\, \partial\, T\, \psi(e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) -
	h\, T\, \psi(\partial\, e_{2m}) \big) \\
	& =
	T\, \psi(e_{2m}) -
	h\, \psi(T\,N\, e_{2m-1}) \big) \\
	& =
	T \psi(e_{2m}).
	\end{align*}
	For $n = 2m$ the proof is analogous.
	The chain map $\psi$ is a quasi-isomorphism since both complexes have the homology of a point and $\psi(e_0) = (1, \dots, r)$ is a generator of the homology.
\end{proof}

@@

We will define a May--Steenrod structure $\psi_{U(\mathcal M)} \colon \mathcal W \to U(\mathcal M)$ by composing the May--Steenrod structure $\psi_{\mathcal X} \colon \mathcal W \to \mathcal X$ with a morphism $SL \colon \mathcal X \to U(\mathcal M)$ of $\mathrm{C}$-modules ($\mathrm{S}$-modules in fact) that we now define.
Given a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\}$ let $SL(s)$ be the element represented by the immersed connected $(1,r)$-graph:

### 2021-08-30 sec/groups.tex
Improvements after Greg's proofreading

Fixing notation, let $\mathrm{S}_r$ be the symmetric group of $r$ elements and $\mathrm{C}_r$ be the cyclic group of order $n$ thought of as the subgroup of $\mathrm{S}_r$ generated by an element~$\rho$.
We denote this inclusion by $\iota \colon \mathrm C_r \to \mathrm S_r$.

@@

Let $\mathrm G$ be a group and $M$ an $R[\mathrm G]$-module, we define the \textit{homology of $\mathrm G$ with coefficients in $M$}, denoted $H(\mathrm G; M)$, as the homology of any chain complex $P \otimes_{R[\mathrm G]} M$ with $P \to R$ a resolution.
We will be particularly interested in the case when $M = \mathbb F_p(q)$ is the trivial or sign $\mathbb F_p[\mathrm{S}_r]$-module depending on the parity of~$q$.

@@

The homology of $\mathrm{S}_r$ is harder to compute.
With untwisted coefficients, the method of computation followed by several authors was to prove the injectivity of this homology into that of the infinite symmetric group and take advantage of a natural Hopf algebra structure on it.
A powerful result stemming from the deep connection of this question with infinite loop space theory, is the existence of a homology isomorphism of spaces
\begin{equation*}
\Z \times B\mathrm{S}_\infty \to Q(S^0) = \Omega^\infty \Sigma^\infty (S^0)
\end{equation*}
credited to Dyer--Lashof \cite{dyer62lashof}, Barratt--Priddy and Quillen \cite{barratt1972priddyquillen}.

In this work, we are interested in the mod $p$ homology of $\mathrm{S}_p$ ($p$ a prime) which, as explained in \cite[Corollary~VI.1.4]{adem2004milgram}, is detected by the inclusion $\iota \colon \mathrm{C}_p \to \mathrm{S}_p$, i.e., $\iota$ induces a surjection in mod $p$ homology.
We now describe the kernel of this surjection.

>>>>

Fixing notation, let $\mathrm{S}_r$ be the symmetric group of $r$ elements and $\mathrm{C}_r$ be the cyclic group of order $r$ thought of as the subgroup of $\mathrm{S}_r$ generated by an element~$\rho$.
We denote this inclusion by $\iota \colon \mathrm C_r \to \mathrm S_r$.

@@

Let $\mathrm G$ be a group and $M$ an $R[\mathrm G]$-module, we define the \textit{homology of $\mathrm G$ with coefficients in $M$}, denoted $H(\mathrm G; M)$, as the homology of any chain complex $P \otimes_{R[\mathrm G]} M$ with $P \to R$ a resolution.
We will be particularly interested in the case when $M = \mathbb F_p(q)$ is the trivial or sign $\mathbb F_p[\mathrm{S}_r]$-module depending on if the parity of~$q$ is even or odd respectively.

@@

The homology of $\mathrm{S}_r$ is harder to compute.
With untwisted coefficients, the method of computation followed by several authors was to prove the injectivity of this homology into that of the infinite symmetric group and take advantage of a natural Hopf algebra structure on it.
A powerful result stemming from the deep connection of this question with infinite loop space theory is the existence of a homology isomorphism of spaces
\begin{equation*}
\Z \times B\mathrm{S}_\infty \to Q(S^0) = \Omega^\infty \Sigma^\infty (S^0)
\end{equation*}
credited to Dyer--Lashof \cite{dyer62lashof}, Barratt--Priddy and Quillen \cite{barratt1972priddyquillen}.

In this work, we are interested in the mod $p$ homology of $\mathrm{S}_p$ ($p$ a prime) which, as explained in \cite[Corollary~VI.1.4]{adem2004milgram}, is detected by a group inclusion $\iota \colon \mathrm{C}_p \to \mathrm{S}_p$, that is, the map induced in mod $p$ group homology is a surjection.
We now describe the kernel of this surjection.

### 2021-08-30 sec/introduction.tex
Improvements after Greg's proofreading

The role Steenrod operations play in stable homotopy theory is hard to overstate.
The reason is that, given the representability of the cohomology functor, these operations together with the Bockstein homomorphism can be used to give a complete description of the algebraic structure naturally present on the mop-$p$ cohomology algebra of spaces.
For the even prime, Steenrod squares were introduced in \cite{steenrod1947products} via an explicit choice of coherent homotopical corrections to the broken symmetry of Alexander--Whitney's chain approximation to the diagonal, the so called cup-$i$ products.
Later, for odd primes, their definition was given non-effectively using arguments based on the mod $p$ homology of symmetric groups \cite{steenrod1952reduced, steenrod1962cohomology, steenrod1962cohomology}.
This viewpoint enhanced the conceptual understanding of the operations and allowed for many advances \cite{adem1952iteration, milnor1958dual, adams1995stable}, but lacked the concreteness of their definition at the even prime.
The purpose of this paper is to fill this gap in the literature, introducing effective descriptions of multioperations at the cochain level -- generalizations of the Steenrod's cup-$i$ products -- that define Steenrod operations at all primes.

@@

In applied topology, the use of persistence homology \cite{edelsbrunner2002topological, carlsson2005barcode} has open many new interdisciplinary research directions \cite{chan2013viral, hess2017cliques}, and the availability of formulae for the cup-$i$ products allowed for the development of a theory of persistence Steenrod modules accessing finer features of the data \cite{medina2018persistence}.

>>>>

The role Steenrod operations play in stable homotopy theory is hard to overstate.
The reason is that, given the representability of the cohomology functor, these operations together with the Bockstein homomorphism can be used to give a complete description of the algebraic structure naturally present on the mop-$p$ cohomology algebra of spaces.
For the even prime, Steenrod squares were introduced in \cite{steenrod1947products} via an explicit choice of coherent homotopical corrections to the broken symmetry of Alexander--Whitney's chain approximation to the diagonal, the so called cup-$i$ products.
Later, for odd primes, their definition was given non-effectively using existence arguments based on the mod $p$ homology of symmetric groups \cite{steenrod1952reduced, steenrod1962cohomology, steenrod1962cohomology}.
This viewpoint enhanced the conceptual understanding of the operations and allowed for many advances \cite{adem1952iteration, milnor1958dual, adams1995stable}, but lacked the concreteness of their definition at the even prime.
The purpose of this paper is to fill this gap in the literature, introducing effective descriptions of multioperations at the cochain level -- generalizations of the Steenrod's cup-$i$ products -- that define Steenrod operations at all primes.

@@

In applied topology, the use of persistence homology \cite{edelsbrunner2002topological, carlsson2005barcode} has created many interdisciplinary research directions \cite{chan2013viral, hess2017cliques}, and the availability of formulae for the cup-$i$ products allowed for the development of a theory of persistence Steenrod modules accessing finer features of the data \cite{medina2018persistence}.
